### 2015-07-06 narrative/qos.tex
more refs fixed
some quick edits.

\paragraph{QoS Background:} 
Quality of Service (QoS) refers to the
properties of the performance quality of a particular service, in
this case the storage system.
%Performance quality can be expressed in either relative or absolute
%terms. Examples of relative performance terms are ``fair'',
%``proportional'', or ``priority/class-based'' (``the higher the
%priority the better the service'' or ``1st class is better than
%2nd''). The key advantage of relative performance terms is that
%they are easy to implement. The key disadvantage is the difficulty
%of making any guarantees based on those implementations --
%even relative guarantees are mired by the well-known effect of
%priority inversion~\cite{lampson:cacm80}. Examples of absolute
%performance terms are ``rate-based'', ``soft real-time'', and
%``hard real-time''. The key advantage of absolute performance
%terms are that they enable the implementation of strong
%guarantees: absolute terms enable agreements between application
%tasks and storage that we call \emph{reservations} and that do not
%change meaning depending
%on the storage system's workload. The key challenge of absolute
%performance terms is that their implementation is non-trivial,
%especially in large-scale systems.
One of the key differentiating features this project offers is driving storage
decisions through QoS \emph{reservations}, rather than continuing the current model
of allowing applications to compete freely without restriction, for part or all of the storage
resources. This traditional approach leads to intereference effects
\cite{lofstead:2010:io-variability,liu_hotstorage} that can greatly impact IO performance predictability. 
%Further, with the introduction of additional tiers in the
%storage stack, the performance variability will increase with greater
%competiion for limited high performance resources. 
%% Temporaily commented out the above sentence, because it may not be true for the next DOE systems. 
Unfettered, this competition can yield a lower system and application performance 
than the case where applications attempt to reserve storage resources, thereby enabling the storage system
to reason over its performance in terms of reservations and actual
performance.
If the system uses reservations intelligently, its performance is predictable and
efficient because resources are not wasted on interference.
In the cases where an appication arrives that needs to run more urgently
than any of the other already running applications, the system
can rapidly respond to it by reducing the amount of resources taken 
by less urgent applications, allowing the urgent application to finish,
and later reverting back to the previous committed resource level for the rest. This can
be automatically enforced via user- (or adiminstrator-) specified \textit{policies}
(section \ref{}).

There are two fundamental causes that can impede the I/O performance
in an extreme-scale SSIO system. One is indirection, where layers
of storage tiers are employed either for performance and/or for
scalability reasons.  The direct consequence is that there could
be multiple traversing paths from application end to the rest place
of data, and more often than not, the I/O paths are not under control
under any single authority. The other cause is the shared use of
resources. The best effort I/O request/response nature and lack of
QoS mechanisms imply that there is little guarantee in terms of
expected performance.  Both indirection and shared use of resource
contribute to a high probability of imbalanced use of resources
thus the occurrence of congestion and degraded performance.
Traditionally, there is a disconnection between storage infrastructure
and rest of system software and applications: the infrastructure
details are not exposed anywhere at all. One argument favoring this
is to ensure a platform agnostic design.

@@

\paragraph{Related Work:} On current HPC systems, absolute IO performance
guarantees are not possible. Previous work \cite{lofstead:2010:io-variability,liu_hotstorage} 
from this project team uses a client-assisted approach to mitigate IO variablity.
Existing efforts on scheduing resources~\cite{io-cop,dorier} attempt to offer
admission control to maximize storage performance. Unfortunately, these efforts
are limited to participating applications and only applications from a single
platform. For this approach to be effective, the storage system must offer an
approach that handles clients from all connected platforms and does not
require modification to use new storage access APIs. By using a priority
tagging system, we will offer defaults suitable for all applications that can
be informed by simple extensions to the job scheduler or more advanced
management through new APIs.

>>>>

\paragraph{QoS Background:} 
Quality of Service (QoS) refers to the
properties of the performance quality of a particular service, in
this case the storage system.
%Performance quality can be expressed in either relative or absolute
%terms. Examples of relative performance terms are ``fair'',
%``proportional'', or ``priority/class-based'' (``the higher the
%priority the better the service'' or ``1st class is better than
%2nd''). The key advantage of relative performance terms is that
%they are easy to implement. The key disadvantage is the difficulty
%of making any guarantees based on those implementations --
%even relative guarantees are mired by the well-known effect of
%priority inversion~\cite{lampson:cacm80}. Examples of absolute
%performance terms are ``rate-based'', ``soft real-time'', and
%``hard real-time''. The key advantage of absolute performance
%terms are that they enable the implementation of strong
%guarantees: absolute terms enable agreements between application
%tasks and storage that we call \emph{reservations} and that do not
%change meaning depending
%on the storage system's workload. The key challenge of absolute
%performance terms is that their implementation is non-trivial,
%especially in large-scale systems.
One of the key differentiating features this project offers is driving storage
decisions through QoS \emph{reservations}, rather than continuing the current model
of allowing applications to compete freely without restriction, for part or all of the storage
resources. This traditional approach leads to intereference effects
\cite{lofstead:2010:io-variability,liu_hotstorage} that can greatly impact IO performance predictability. 
%Further, with the introduction of additional tiers in the
%storage stack, the performance variability will increase with greater
%competiion for limited high performance resources. 
%% Temporaily commented out the above sentence, because it may not be true for the next DOE systems. 
Unfettered, this competition can yield a lower system and application performance 
than the case where applications attempt to reserve storage resources, thereby enabling the storage system
to reason over its performance in terms of reservations and actual
performance.
In the cases where an appication arrives that needs to run more urgently
than any of the other already running applications, the system
can rapidly respond to it by reducing the amount of resources taken 
by less urgent applications, allowing the urgent application to finish,
and later reverting back to the previous committed resource level for the rest.
This can be automatically enforced via user- (or adiminstrator-) specified \textit{policies}
(section \ref{}).

There are two fundamental causes that can impede the I/O performance
in an extreme-scale SSIO system. One is indirection, where layers
of storage tiers are employed either for performance and/or for
scalability reasons.  The direct consequence is that there could
be multiple traversing paths from application end to the rest place
of data, and more often than not, the I/O paths are not under control
under any single authority. The other cause is the shared use of
resources. The best effort I/O request/response nature and lack of
QoS mechanisms imply that there is little guarantee in terms of
expected performance. Both indirection and shared use of resource
contribute to a high probability of imbalanced use of resources
thus the occurrence of congestion and degraded performance.

@@

\paragraph{Related Work:} On current HPC systems, absolute IO performance
guarantees are not possible. Previous work \cite{lofstead:2010:io-variability,liu_hotstorage} 
from this project team uses a client-assisted approach to mitigate IO variablity.
Existing efforts on scheduing resources~\cite{thapaliya:2014:io-cop,dorier:2014:calciom} attempt to offer
admission control to maximize storage performance. Unfortunately, these efforts
are limited to participating applications and only applications from a single
platform. For this approach to be effective, the storage system must offer an
approach that handles clients from all connected platforms and does not
require modification to use new storage access APIs. By using a priority
tagging system, we will offer defaults suitable for all applications that can
be informed by simple extensions to the job scheduler or more advanced
management through new APIs.
